---
title: "Untitled"
author: "tata iles - saidani akram"
date: "2023-11-20"
output:
  html_document:
    toc: true
    theme: united
    fontsize: 12pt
---

\tableofcontents

```{r include=FALSE}
library(tidyverse)
library(highcharter)
library(vars)
library(lmtest)
library(viridis)
library(forecast)
library(tseries)
library(ARDL)
```

```{r include=FALSE}
data <- read.csv("D:/Python-R/Databases/Données Model ardl.csv")
```

# Data Overview

```{r echo=FALSE}
knitr::kable(head(data), col.names = gsub("[.]", " ", names(data)),caption = "DATA")
```

# Series Visualization

```{r echo=FALSE}
data_st <- cbind(data$year,scale(data[,2:4],center = TRUE,scale = TRUE)) %>% as.data.frame()

highchart() %>%
  hc_colors(viridis(3)) %>% 
  hc_title(text = "Series Graphs") %>%
  hc_xAxis(type = "datetime") %>%
  hc_yAxis(title = list(text = "Values")) %>%
  hc_add_series(data,"line", hcaes(x = data_st$V1, y = data_st$gdp), name = "GDP") %>%
  hc_add_series(data,"line", hcaes(x = data_st$V1, y = data_st$energy), name = "Energy") %>%
  hc_add_series(data,"line", hcaes(x = data_st$V1, y = data_st$train), name = "Train")
```

# Modeling

## Causality test

No causality effects from the dependent variable

```{r echo=FALSE}
cat("Does gdp cause energy \n ")
grangertest(data$gdp,data$energy)
cat("----------------------------------------------------\nDoes gdp cause train\n")
grangertest(data$gdp,data$train)
```

## Stationarity tests

```{r echo=FALSE}
ggtsdisplay(data$gdp)
ggtsdisplay(data$energy)
ggtsdisplay(data$train)
```

```{r echo=FALSE}
apply(data[,2:4],2,pp.test)
```

## Stationarity Results

```{r echo=FALSE}
statio <- data.frame("Serie"=colnames(data[,2:4]),"pp.test"=c("I(1)","I(1)","I(1)"))
knitr::kable(statio, col.names = gsub("[.]", " ", names(statio)))
```

## Select ARDL Order

```{r echo=FALSE, warning=FALSE}
data_ts <- ts(data = data[, -1], start = 1920,end=1957, frequency = 1)

AIC_selection <- auto_ardl(data = data_ts, max_order = 5, selection = "AIC",

formula = gdp ~ energy + train | crisis29,
selection_minmax = "min", search_type = "horizontal",
start = 1920, end = 1957, grid = TRUE)

AIC_selection
```

## Best Model

```{r echo=FALSE}
ardl_model<- ardl(data = data_ts, order = AIC_selection$best_order,

formula = gdp ~ energy + train | crisis29, start = 1920, end = 1957)

summary(ardl_model)
```

## Validation tests

### Autocorrelation & Heteroscedasticity

Residuals are not autocorrelated and are homoscedastic

```{r echo=FALSE}
for (i in 1:2){
godfrey<-bgtest(ardl_model, order = i, type = "Chisq")
print(godfrey)
}
print(bptest(ardl_model, studentize = TRUE))
```

### Residual plot & Normality 

```{r echo=FALSE}
ardl_model %>% checkresiduals()
```

### Model Stability

#### Short term

Model is stable

```{r echo=FALSE}
for (i in 2:3){
resettestt<-resettest(ardl_model, power = 2:i, type = "fitted")
print(resettestt)
}
```

#### Long term

Model is stable

```{r echo=FALSE}
f_bound <- bounds_f_test(ardl_model, case = 3, exact = TRUE, R = 40000, test = "F",vcov_matrix = NULL)
f_bound
```

## Final Equation

### Short term

```{r echo=FALSE}
long_run <- multipliers(ardl_model, type = "lr", vcov_matrix = NULL)
print(long_run)
```

### Long term

```{r echo=FALSE}
restricted<- recm(ardl_model, case = 3)
summary(restricted)
```
