---
title: "MPRA"
author: ""
date: "2023-10-05"
output: html_document
---

```{r include=FALSE}
library(tidyverse)
library(fpp)
library(forecast)
library(tseries)
library(readxl)
library(vars)
library(urca)
library(lmtest)
library(ggforce)
library(tsDyn)
library(TSstudio)
library(moments)
library(skimr)
library(jtools)
```

# DATA

```{r include=FALSE}
gdp <- read.csv("D:/Python-R/Databases/mpra/realgdp/c7d1e6d6-ab6c-4740-907f-1194646bec9e_Data.csv")

hfce <- read.csv("D:/Python-R/Databases/mpra/hfce/d93ca812-b285-4f8b-b30f-8b688872f28a_Data.csv")

i <- read.csv("D:/Python-R/Databases/mpra/realinterstrate/d25ecd02-8d03-4083-968d-4a91941dcffc_Data.csv")
```

```{r include=FALSE}
gdp <- t(gdp) %>% as.data.frame() %>% slice(5:38) 

hfce <- t(hfce) %>% as.data.frame() %>%  slice(5:38)

i <- t(i) %>% as.data.frame() %>%  slice(5:38)
```

```{r include=FALSE}
nigeria <- cbind(gdp$V2,hfce$V1,i$V1)%>%as.data.frame()
soaf <- cbind(gdp$V1,hfce$V2,i$V2)%>%as.data.frame()
```

```{r include=FALSE}
colnames(nigeria) <- c("gdp","hfce","i%")
colnames(soaf) <- c("gdp","hfce","i%")
```

```{r include=FALSE}
nigeria <- nigeria %>%
  mutate_if(is.character,as.numeric) %>% mutate(gdp = log(gdp),
                                                hfce = log(hfce))

soaf <- soaf %>%
  mutate_if(is.character,as.numeric) %>% mutate(gdp = log(gdp),
                                                hfce = log(hfce))
```

```{r include=FALSE}
median(nigeria$hfce,na.rm = TRUE) -> med

nigeria$hfce <- replace_na(nigeria$hfce,med)
```

```{r echo=FALSE}
nigsh <- nigeria %>% mutate(Year = 1980:2013)
sosh <- soaf %>% mutate(Year = 1980:2013)
knitr::kable(head(nigsh), col.names = gsub("[.]", " ", names(nigsh)),caption = "Nigeria")
knitr::kable(head(sosh), col.names = gsub("[.]", " ", names(sosh)),caption = "South Africa")
```

# Summary Stats

## NIGERIA

```{r echo=FALSE, warning=FALSE}
summary_nigeria <- nigeria %>%
  summarise_all(list(
    Mean = ~ mean(.),
    Median = ~ median(.),
    Max = ~ max(.),
    Min = ~ min(.),
    std = ~ sd(.),
    Skewness = ~ skewness(.),
    Kurtosis = ~ kurtosis(.),
    Jarque_Bera = ~ jarque.test(.)$statistic
  )) %>%
  gather(key, value) %>%
  separate(key, into = c("Statistic", "Variable"), sep = "_") %>%
  spread(Statistic, value) %>% mutate_if(is.numeric,round,3)

knitr::kable(summary_nigeria, col.names = gsub("[.]", " ", names(summary_nigeria)))
```

## South Africa

```{r echo=FALSE, warning=FALSE}
summary_soaf <- soaf %>%
  summarise_all(list(
    Mean = ~ mean(., na.rm = TRUE),
    Median = ~ median(., na.rm = TRUE),
    Max = ~ max(., na.rm = TRUE),
    Min = ~ min(., na.rm = TRUE),
    std = ~ sd(.,na.rm = TRUE),
    Skewness = ~ skewness(., na.rm = TRUE),
    Kurtosis = ~ kurtosis(., na.rm = TRUE),
    Jarque_Bera = ~ jarque.test(.)$statistic
  )) %>%
  gather(key, value) %>%
  separate(key, into = c("Statistic", "Variable"), sep = "_") %>%
  spread(Statistic, value) %>% mutate_if(is.numeric,round,3)

knitr::kable(summary_soaf, col.names = gsub("[.]", " ", names(summary_soaf)))
```

# Stationnarity

## Graphes (acf/pacf)

### Nigeria

```{r echo=FALSE}
nigeria$gdp %>% ggtsdisplay(main="Nigerian Real gdp")
nigeria$hfce %>% ggtsdisplay(main="Nigerian hfce")
nigeria$`i%` %>% ggtsdisplay(main="Nigerian Real Interest Rate")
```

### South Africa

```{r echo=FALSE}
soaf$gdp %>% ggtsdisplay(main="South Africa's Real gdp")
soaf$hfce %>% ggtsdisplay(main="South Africa's hfce")
soaf$`i%` %>% ggtsdisplay(main="South Africa's Real Interst Rate")
```

From the graphs we can see that

```{r echo=FALSE}
adf <- data.frame("Country"=c("Nigeria","South Africa"),
                       "GDP"=c("not stationary","not stationary"),
                       "HFCE"=c("not stationary","not stationary"),
                       "Interest Rate" = c("stationary","stationary"))
knitr::kable(adf, col.names = gsub("[.]", " ", names(adf)))
```

## ADF test

### Nigeria

```{r echo=FALSE}
apply(nigeria,2,adf.test)
```

```{r include=FALSE}
cat("GDP\n\n")
ur.df(nigeria$gdp,selectlags="AIC",type="drift")%>%summary()
cat("nHFCE\n\n")
ur.df(nigeria$hfce,selectlags="AIC",type="drift")%>%summary()
cat("nRIR\n\n")
ur.df(nigeria$`i%`,selectlags="AIC",type="drift")%>%summary() 
```

### South Africa

```{r echo=FALSE}
apply(soaf,2,adf.test)
```

```{r include=FALSE}
cat("GDP\n\n")
ur.df(soaf$gdp,selectlags="AIC",type="drift")%>%summary()
cat("HFCE\n\n")
ur.df(soaf$hfce,selectlags="AIC",type="drift")%>%summary()
cat("RIR\n\n")
ur.df(soaf$`i%`,selectlags="AIC",type="drift")%>%summary() 
```

## Stationarizing

```{r echo=FALSE}
diff(nigeria$gdp) %>% ggtsdisplay(main="Nigerian Real gdp")
nigeria$hfce %>% diff() %>% ggtsdisplay(main="Nigerian hfce")
```

```{r echo=FALSE}
adf.test(diff(nigeria$`i%`))
```

```{r echo=FALSE}
diff(soaf$gdp) %>% ggtsdisplay(main="South africa's Real gdp")
diff(soaf$hfce) %>% diff() %>% ggtsdisplay(main="South africa's hfce")
diff(soaf$`i%`) %>% diff() %>% ggtsdisplay(main="South africa's hfce")
```

```{r echo=FALSE, warning=FALSE}
adf.test(diff(soaf$`i%`))
```

From the test we see that non of the series is stationary and taking the first difference makes them stationary

```{r echo=FALSE}
adf <- data.frame("Country"=c("Nigeria","South Africa"),
                       "GDP"=c("I(1)","I(1)"),
                       "HFCE"=c("I(1)","I(1)"),
                       "Interest Rate" = c("I(1)","I(1)"))
knitr::kable(adf, col.names = gsub("[.]", " ", names(adf)))
```

# Modelisation ECM

```{r warning=FALSE, include=FALSE}
VARselect(cbind(ts(nigeria$gdp),ts(nigeria$hfce),ts(nigeria$`i%`)),lag.max = 8,type = "const")[["selection"]]

          
VARselect(cbind(ts(soaf$gdp),ts(soaf$hfce),ts(soaf$`i%`)),lag.max = 8,type = "const")[["selection"]]
```

## Granger causality and Johanson cointegration tests

### Nigeria

```{r echo=FALSE, warning=FALSE}
diff_nigeria <- cbind(diff(nigeria$gdp),diff(nigeria$hfce),diff(nigeria$`i%`)) 

var=VAR(cbind(diff(nigeria$gdp),diff(nigeria$hfce),diff(nigeria$`i%`)),
        p=5,type = "const")

causality(var,cause = c("y1","y2"))   
causality(var,cause = c("y2","y3"))
causality(var,cause = c("y1","y3"))
```

```{r echo=FALSE}
ca.jo(cbind(ts(nigeria$gdp),ts(nigeria$hfce),ts(nigeria$`i%`)),type = "trace",ecdet = "const",K=6)%>%summary()
```

### South Africa

```{r echo=FALSE, warning=FALSE}
diff_saof <- cbind(diff(soaf$gdp),diff(soaf$hfce),diff(soaf$`i%`)) 

var2=VAR(cbind(diff(soaf$gdp),diff(soaf$hfce),diff(soaf$`i%`)),
        p=5,type = "const")

causality(var2,cause = c("y1","y2"))   
causality(var2,cause = c("y2","y3"))
causality(var2,cause = c("y1","y3"))
```

```{r echo=FALSE, paged.print=TRUE}
ca.jo(cbind(ts(soaf$gdp),ts(soaf$hfce),ts(soaf$`i%`)),type = "trace",ecdet = "const",K=2)%>%summary()
```

```{r echo=FALSE}
granger <- data.frame("Country"=c("Nigeria","South Africa"),"Result"=c("gdp interst rate cause hfce","gdp interst rate cause hfce"))
knitr::kable(granger, col.names = gsub("[.]", " ", names(granger)),caption = "Granger test results")
```

```{r echo=FALSE}
johanson <- data.frame("Country"=c("Nigeria","South Africa"),"Result"=c("contegrated (1)","not cointegrated"))
knitr::kable(johanson, col.names = gsub("[.]", " ", names(johanson)),caption = "Johanson test results")
```

# Regresssion

$HFCE_t = Ω + Ω_1lnRGDP_t + Ω_2lnHFCE_{t-1} + Ω_3RIR_t + η_t$

## Nigeria

```{r include=FALSE}
ln_HFCE1 = nigeria$hfce
ln_RGDP1 = nigeria$gdp
RIR1 =nigeria$`i%`

ln_HFCE1_t_1 = lag(ln_HFCE1,1)

reg_1_data <- cbind(ln_HFCE1,ln_RGDP1,ln_HFCE1_t_1,RIR1) %>% as.data.frame()
```

```{r include=FALSE}
model <- lm(ln_HFCE1 ~ ln_RGDP1+ln_HFCE1_t_1+RIR1, data = reg_1_data)


residuals <- residuals(model)


lagged_residuals <- lag(residuals, 1)

reg_1_data$`AR(1)` <- c(NA,lagged_residuals)
```

```{python}
import statsmodels.api as sm
```

```{python}
data1 = r.reg_1_data

X1 = data1[["ln_RGDP1","ln_HFCE1_t_1","RIR1"]]
Y1 = data1[["ln_HFCE1"]]

X1=sm.add_constant(X1)
result1=sm.OLS(Y1,X1,missing= "drop").fit()
print(result1.summary())
```

$HFCE_t = -1.2758 + 0.2921lnRGDP_t + 0.7533lnHFCE_{t-1} + 0.0094RIR_t + η_t$

## South Africa

```{r include=FALSE}
ln_HFCE2 = soaf$hfce
ln_RGDP2 = soaf$gdp
RIR2 =soaf$`i%`

ln_HFCE2_t_1 = lag(ln_HFCE2,1)

reg_2_data <- cbind(ln_HFCE2,ln_RGDP2,ln_HFCE2_t_1,RIR2) %>% as.data.frame()
```

```{python}
data2 = r.reg_2_data

X2 = data2[["ln_RGDP2","ln_HFCE2_t_1","RIR2"]]
Y2 = data2[["ln_HFCE2"]]

X2=sm.add_constant(X2)
result2=sm.OLS(Y2,X2,missing= "drop").fit()
print(result2.summary())
```

$HFCE_t = -1.318 + 0.9921lnRGDP_t + 0.0409lnHFCE_{t-1} + 0.00375RIR_t + η_t$

Adaptive expectation coefficient for Nigeria 𝝍 =1 - .753369 = 0.246631

Adaptive expectation coefficient for South Africa 𝝍 =1- 0.040977 = 0.959023

```{r include=FALSE}
final <- data.frame("Country"=c("Nigeria","South Africa"),
           "Elasticity of Consumption with respect to actual income"=c(0.292169,0.992103),
           "Adaptive expectation coefficient(λ)"=c( 0.246631,0.959023),
           "Elasticity of consumption with Respect to permanent income(β)"=c(0.292169/0.246631,0.992103/0.959023))
```

```{r echo=FALSE}
knitr::kable(final, col.names = gsub("[.]", " ", names(final)))
```

### Adding the AR(1) for Nigeria regression

```{python}
data1 = r.reg_1_data

X1 = data1[["ln_RGDP1","ln_HFCE1_t_1","RIR1","AR(1)"]]
Y1 = data1[["ln_HFCE1"]]

X1=sm.add_constant(X1)
result1=sm.OLS(Y1,X1,missing= "drop").fit()
print(result1.summary())

```
